Predicting Survival Probability of NASA Aircraft Engines

### 2025-04-22 slides.qmd
Final slides update 4/22

::: incremental
-   A statistical regression method specializing in modeling
    time-to-event predictions with survival data [@Abeysekera2009]

@@

-   Primarily used in the health field but has applications in
    predicting bank failure, the survival probability of machines, and
    insurance likelihood payouts
:::

@@

::: incremental
-   The model assumes that as time goes on, the survival probability
    will approach zero with no survivors [@asghar2024]

@@

-   The model cannot provide a specific value for when the event will
    happen, only the probability of when the event might happen
:::

@@

::: incremental
-   Independence assumption

@@

    -   Assumes the ratio of hazards rates for any two subjects are
        constant at all times [@Bustan2018]
:::

@@

::: incremental
-   Model accuracy evaluation is done using the concordance index

    ::: incremental
    -   The concordance index measures the amount of agreement between
        two variables

    -   A value of 1 means all the pairs are correctly ordered while a
        value of 0 means no pairs are correctly ordered
    :::

@@

    ::: incremental
    -   If the probability is $\geq 50$%, it is assumed the event has
        not occurred

    -   If the probability is $< 50$%, it is assumed the event has
        occurred
    :::
:::

@@

The table provides the model number, the covariates used, the AIC and
BIC from the stepwise regression if applicable, and the concordance
index.

@@

all_cov <- coxph(data = data_grouped, surv_object ~ os1 + os2 + os3 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm13 + sm14 + sm15 + sm16 + sm17 + sm18 + sm19 + sm20 + sm21)

aic <- MASS::stepAIC(all_cov, trace = 0)
n <- nrow(data_grouped)
bic <- step(all_cov, direction = "both", k = log(n), trace = 0)

cox1 <- coxph(data = data_grouped, surv_object ~ os1 + os2 + os3 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm13 + sm14 + sm15 + sm16 + sm17 + sm18 + sm19 + sm20 + sm21)

cox2 <- coxph(data = data_grouped, surv_object ~ os3 + sm3 + sm4 + sm5 + sm8 + sm9 + sm15 + sm16 + sm17 + sm18 + sm20)

cox3 <- coxph(data = data_grouped, surv_object ~ os3 + sm3 + sm4 + sm8 + sm9 + sm15 + sm18)

cox4 <- coxph(data = data_grouped, surv_object ~ os3 + sm13 + sm14 + sm19)

stepwise_results <- data.frame(c("model 1", "model 2", "model 3", "model 4"), 
                               c("All covariates", "os3, sm3 - sm5, sm8 - sm9, sm15 - sm18, sm20", "os3, sm3 - sm4, sm8-sm9, sm15, sm18", "os3, sm13 - sm14, sm19"),
                               c("2449.92", "2431.88", "N/A", "N/A"),
                               c("2547.71", "N/A", "2461.31", "N/A"),
                               c(cox1$concordance[6], cox2$concordance[6], cox3$concordance[6], cox4$concordance[6]))

colnames(stepwise_results) <- c("Model", "Covariates", "AIC", "BIC", "Concordance")

kable(stepwise_results, format = "markdown")
```

Model for the continuing analysis will have all covariates except sm16
and sm19.

@@

-   Based on the below plots, this assumption is met.

@@

-   Based on the below table and plots, this assumption is met

>>>>

-   A statistical regression method specializing in modeling
    time-to-event predictions with survival data [@Abeysekera2009]

@@

-   Primarily used in the health field but has applications in
    predicting bank failure, the survival probability of machines, and
    insurance likelihood payouts

@@

-   The model assumes that as time goes on, the survival probability
    will approach zero with no survivors [@asghar2024]

@@

-   The model cannot provide a specific value for when the event will
    happen, only the probability of when the event might happen

@@

-   Independence assumption

@@

    -   Assumes the ratio of hazards rates for any two subjects are
        constant at all times [@Bustan2018]

@@

-   Model accuracy evaluation is done using the concordance index

    -   The concordance index measures the amount of agreement between
        two variables

    -   A value of 1 means all the pairs are correctly ordered while a
        value of 0 means no pairs are correctly ordered

@@

    -   If the probability is $\geq 50$%, it is assumed the event has
        not occurred

    -   If the probability is $< 50$%, it is assumed the event has
        occurred

@@

-   After an analysis of the covariates provided with the data, we
    determined that sm16 and sm19 should not be included in the model

    -   sm16 produces an infinite coefficient when taking the $log$

    -   sm19 is linearly dependent on another covariate

$$
\begin{split}
h(t) = h_0(t)\text{exp}(&-3.97\text{ os1}+26.06\text{ os2}+121.38\text{ os3}+30.47\text{ sm1} \\ &+0.06\text{ sm2}+0.04\text{ sm3}+0.04\text{ sm4} + 48.62\text{ sm5}\\ &-39.93\text{ sm6}-0.10\text{ sm7}+2.26\text{ sm8}-0.03\text{ sm9} \\ &+15.22\text{ sm10} +0.43\text{ sm11}+0.01\text{ sm12}-1.38\text{ sm13}\\ &+0.02\text{ sm14}+8.75\text{ sm15}+0.11\text{ sm17} -14.86\text{ sm18} \\ &-1.02\text{ sm20} -0.34\text{ sm21})
\end{split}
$$

@@

cox1.1 <- coxph(data = data_grouped, surv_object ~ os1 + os2 + os3 + sm1 + sm2 + sm3 + sm4 + sm5 + sm6 + sm7 + sm8 + sm9 + sm10 + sm11 + sm12 + sm13 + sm14 + sm15 + sm17 + sm18 + sm20 + sm21)

```

@@

@@
